### 2017-06-16 manuscript/manuscript.Rmd
Add basics of how to use command line

It is important to understand that unlike an operating system's (OS) default file viewer, such as Finder (Apple computers) or File Explorer (?? Windows), Git[^git-name] is not a standalone program for navigating files and folders on a computer. Instead, it adds functionality to the OSs existing file system, by making available a specific set of functions--either executed from the command line, or through a graphical user interface (GUI) / integrated development environment (IDE)--that allow taking snapshots of the project and its files, and distributing the work across multiple computers and users. To begin using Git, users must first download and install the Git software on their computers. Git is free and open source, works on Windows, Macintosh, and Linux OSs (among others). It can be freely downloaded at <https://git-scm.com/>. Before we explain the basics of using Git, we present brief installation instructions for Windows and Apple users.

@@

Even if you already have Git installed (some computers do) it is a good idea to install the latest version[^git-version], which can be downloaded as a standalone program from <https://git-scm.com/download>. Here, we provide more detailed instructions on installing Git for OS X and Windows, but Linux (and other OS) users can find instructions at the Git website (<https://git-scm.com/book/en/v2/Getting-Started-Installing-Git>). After installation instructions, we detail how to configure the Git program so that it is ready to be used.

### OS X

Many OS X computers already have Git installed, especially those operating OS X 10.9 or higher, but it is good practice to install the latest version. The easiest way to update Git to the latest version is to download the installer from <http://git-scm.com/download/mac>, and install it like any other program. Once Git is installed, it can be used from the command line (the OS X command line is available through the Terminal app, which is included by default with the OS X install), or through various GUIs. Although it might at first appear intimidating because of the archaic-looking interface, we encourage using Git from the command line.

### Windows

@@

To operate properly, Git needs to be able to identify its user. You can set these by entering a few basic commands in the command line. To verify the current user information, type `git config --global user.name`, and hit return. This should not return anything, unless a previous user of the computer has set the global Git user name.

To ensure that Git knows who you are, type `git config --global user.name "User Name"` (where User Name is your name) in the command line, and hit return. If you now re-run the first command, the command line will return the name you entered as `"User Name"`. The username can be anything you'd like, but it is probably a good idea to use your real name so that potential collaborators know who you are. The second piece of information is your email, which can be entered by `git config --global user.email "email@address.com"` (where email@address.com is your email address). You can verify that the correct email address was saved by typing `git config --global user.email` in the Terminal, and hit return. Once this information is entered, Git will know who you are, and is thus able to track who is doing what within a project, which is especially helpful when you are collaborating with other people, or when you are working on multiple computers.

>>>>

Git[^git-name], unlike the operating system's (OS) default file viewer (Mac's Finder, Windows' File Explorer), is not a standalone program for navigating files and folders on a computer. Instead, Git adds functionality to the existing file system, by making available a specific set of functions--either executed from the command line, or through a graphical user interface (GUI)--that allow keeping track of files and their history, and distributing the files across multiple computers and users. Because Git is a standalone program and not usually included in standard OS installs, users must first download and install the Git software on their computers. Git is free and open source, works on Windows, Mac, and Linux OSs (among others). It can be freely downloaded at <https://git-scm.com/>. Before we explain the basics of using Git, we present brief installation instructions for Windows and Mac users.

@@

Even if you already have Git installed (some computers do) it is good practice to install the latest version[^git-version], which can be downloaded as a standalone program from <https://git-scm.com/download>. Here, we provide more detailed instructions on installing Git for Mac and Windows OSs, but Linux (and other) users can find instructions at the Git website (<https://git-scm.com/book/en/v2/Getting-Started-Installing-Git>). After installation instructions, we detail how to configure the Git program so that you can start using it.

### Installing Git on Macs

Some Mac computers already have Git installed, especially those operating OS X 10.9 or higher, but it is good practice to install the latest version. The easiest way to update Git to the latest version is to download the installer from <http://git-scm.com/download/mac>, and install it like any other program.

### Installing Git on Windows

@@

Once Git is installed, it can be used from the command line or through a GUI--there are various GUIs for using Git, and below we will explain how to use the R Studio Git GUI. First, it is important to learn a few basic Git commands from the OSs command line. The command line is a text-based interface for interacting with your computer, and its functionality greatly extends that of the standard way of interacting with the computer by clicking and pointing with a mouse.

To access the command line interface, you need to use a command line "shell". Mac users can open the built-in app *Terminal*, and Windows users can use one of many alternatives, such as Microsoft PowerShell (<https://msdn.microsoft.com/powershell>). After opening the shell app, you can type in commands to access the interface. For example typing `ls` (hit Return to execute) lists the files in the current folder.

The first step in using Git is making Git aware of who is using the computer. You need to set the user name and email by entering a few basic commands in the command line. First, to show the current user information, type `git config --global user.name`, and hit return. This should not return anything, unless a previous user of the computer has set the global Git user name.

Each Git command starts with the word `git`, then a command (such as `config`), and then arguments to the command, such as `--global` (for global configuration) followed by the `user.name` variable name to show the global user name. For detailed instructions on how to use Git commands, you can type `git --help`. For help on how to use the `git config` command, type `git config --help`.

To ensure that Git knows who you are, type `git config --global user.name "User Name"` (where User Name is your name) in the command line, and hit return. This command maps `"User Name"` to Git's global `user.name` variable. If you now re-run the first command (`git config --global user.name`), the command line will return the name you entered as `"User Name"`. The user name can be anything you'd like, but it is probably a good idea to use your real name so that potential collaborators know who you are. The second piece of information is your email address, which is entered by `git config --global user.email "email@address.com"` (where `"email@address.com"` is your email address). You can verify that the correct email address was saved by typing `git config --global user.email` in the Terminal, and hit return.

Once this information is entered, Git will know who you are, and is able to track who is doing what and when within a project, which is especially helpful when you are collaborating with other people, or when you are working on multiple computers.

For users who have not previously used text-based command line interfaces, typing out these commands may seem intimidating, but to get started you only need two configuration commands; the basic use of Git requires only 4 commands. If using the text-based command line seems a daunting task, Codecademy (<https://www.codecademy.com/learn/learn-the-command-line>) has a free interactive online tutorial, and MIT has an online game to teach using the command line (<http://web.mit.edu/mprat/Public/web/Terminus/Web/main.html>).
